### 2020-07-08 notebook/bad-forecast-good-decision.Rmd
reknit

---
title: "bad forecast good decision"
author: "Carl Boettiger"
output: github_document
---

```{r graphics_setup, message=FALSE, warning=FALSE}
## Plotting theme settings -- aesthetics only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
#theme_set(theme_solarized(base_size=16))

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(6)
txtcolor <- "#586e75"
```

```{r setup, message=FALSE}
## Required computational libraries
library(tidyverse)
library(MDPtoolbox)
library(expm)
library(mdplearning) # remotes::install_github("boettiger-lab/mdplearning")
```

```{r model_definitions}
states <- seq(0,24, length.out = 240)
actions <- states
obs <- states
sigma_g <- 0.05
reward_fn <- function(x,h) pmin(x,h)
discount <- 0.99

# K is at twice max of f3; 8 * K_3 / 5
f1 <- function(x, h = 0, r = 2, K = 10 * 8 / 5){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}
f2 <- function(x, h = 0, r = 0.5, K = 10){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}

# max is at 4 * K / 5 
f3  <- function(x, h = 0, r = .002, K = 10){
  s <- pmax(x - h, 0)
  s + s ^ 4 * r * (1 - s / K)
}

# looks even closer to f2; not used
#f4  <- function(x, h = 0, r = 4e-7, K = 10){
#  s <- pmax(x - h, 0)
#  s + s ^ 8 * r * (1 - s / K) + s * (1 - s / K)
#}

models <- list(f1 = f1, f2 = f2, f3 = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)

true_model <- "f3"
```

```{r transition_matrices}
# A function to compute the transition matrices for each model:
transition_matrices <- function(f,
                      states,
                      actions,
                      sigma_g){

  n_s <- length(states)
  n_a <- length(actions)

  transition <- array(0, dim = c(n_s, n_s, n_a))
  for (k in 1:n_s) {
    for (i in 1:n_a) {
      nextpop <- f(states[k], actions[i])
      if(nextpop <= 0){
        transition[k, , i] <- c(1, rep(0, n_s - 1))
      } else if(sigma_g > 0){
        x <- dlnorm(states, log(nextpop), sdlog = sigma_g)
        if(sum(x) == 0){ ## nextpop is computationally zero
          transition[k, , i] <- c(1, rep(0, n_s - 1))
        } else {
          x <- x / sum(x) # normalize evenly
          transition[k, , i] <- x
        }
      } else {
        stop("sigma_g not > 0")
      }
      reward[k, i] <- reward_fn(states[k], actions[i])
    }
  }
  transition
}

## Compute reward matrix (shared across all models)
n_s <- length(states)
n_a <- length(actions)
reward <- array(0, dim = c(n_s, n_a))
for (k in 1:n_s) {
  for (i in 1:n_a) {
    reward[k, i] <- reward_fn(states[k], actions[i])
  }
}

## Now compute the transition matrices for each model
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("f1", "f2", "f3")

```

```{r sdp, results="hide"}
## This is the most (only) computationally expensive code chunk:
policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp_value_iteration(P, reward, discount = discount)
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")
```

```{r simulations}
library(mdplearning)
Tmax <- 100
x0 <- which.min(abs(states - 6))
reps <- 100
set.seed(12345)

## Simulate each policy reps times, with `f3` as the true model:

sims <- map_dfr(names(transitions), 
                function(m){
                  policy <- policies %>% filter(model == m) %>% pull(policy)
                  map_dfr(1:reps, 
                          function(i){
                            mdp_planning(transitions[[true_model]], reward, discount,
                                     policy = policy, x0 = x0, Tmax = Tmax) %>%
                              select(value, state_index = state, time, action_index = action)  %>% 
                              mutate(state = states[state_index]) # index->value
                            },
                          .id = "reps")
                },
                .id = "model")

```

## Forecast skill

Note the low variability in the predicted value is a result of the decision strategy, which tweaks harvest up or down with variation.

```{r plot_stepahead}
compare <- sims %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         low = states[ max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[ min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[ next_state],
         model = as.character(model)) %>%
 select(time, model, true, expected, low, high, prob, reps)

compare %>% filter(model != "3") %>%  filter(reps == 3, time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = true), pch = "*", size = 12) +
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high))
```

```{r plot_stepahead_unfished}
compare_unfished <- sims
compare_unfished$state_index <- rep(sims$state_index[sims$model == "1"],3)
compare_unfished <- compare_unfished %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         low = states[ max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[ min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[ next_state ],
         model = as.character(model)) %>% filter(reps == 2, time < 10, model != "3")

compare_unfished %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, col = "grey20", show.legend = FALSE)
```

```{r prob_scores, message = FALSE}
scores <- sims %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  filter(model < 3) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index] ) %>% 
  group_by(reps, model) %>%
  summarise(score = sum(log(prob)))

scores %>% group_by(model) %>% summarise(mean_score = mean(score))

scores %>% mutate(model = as.factor(model)) %>% 
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 70)

```

## Simulations

```{r plot_sims}
fig_ts <- 
  sims %>%
  filter(time < 25, reps < 5) %>%
  ggplot(aes(time, state, col=model, group = interaction(model,reps))) + 
  geom_line(alpha=0.3) 
fig_ts
```

```{r plot_npv, message = FALSE, fig.cap = "Corresponding utility (measured as mean net present value, that is: cumulative value, discounting future values by $\\delta^t$, averaged across replicates). Note that the exepcted utility under model 1, which has the worst forecast, is nearly identical to the optimal utility achieved by managing under the correct model, 3.  The utility derived from model 2 is far smaller, despite it's overall better performance in long term forecasts."}
##  Net Present Value accumulates over time, equivalent for models with near-identical management stategy
npv_df <- sims %>% 
  group_by(model, reps) %>%
  mutate(npv = cumsum(value * discount ^ time)) %>%
  group_by(time, model)  %>% 
  summarise(mean_npv = mean(npv)) %>% 
  arrange(model, time) %>% ungroup()

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")

```

```{r plot_models}
d <- map_dfc(models, function(f) f(states) - states) %>% mutate(state = states)
d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model)) +
  geom_point() + 
  geom_hline(aes(yintercept = 0)) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16))
```

```{r plot_policies}
policies %>%
  ggplot(aes(states,escapement, col=model, lty=model)) + geom_line(lwd=2)
```

## Conclusions

- A good forecast does not mean good management
- A bad forecast does not mean the model is bad for management
- A model permitting successful management does not imply the model is "correct" or generally "good at forecasting"

Model with the egregiously optimistic long-term forecast for the stock size of an unexploited fishery nevertheless actually leads to more conservative management.  In contrast, the model which correctly predicts the long-term average stock size without fishing nevertheless leads to substantial over-harvesting.

```{r}
```

# Appendix

### Unharvested forecasting performance

Here we examine the forecasts under each model, `Tmax` steps into the future, 
ignoring harvesting.  This largely highlights only the differences in the 
unfished equilibria of the models.  It less surprising that bad performance
in predicting the long-term equilibrium would not necessarily mean bad management.

```{r}

sim <- function(f, x0, Tmax, reps = 1){
  map_dfr(1:reps, 
          function(i){
            x <- numeric(length(Tmax))
            x[1] <- x0
            for(t in 2:Tmax){
              mu <- f(x[t-1])
              x[t] <-  rlnorm(1, log(mu), sdlog = sigma_g)
            }
            tibble(t = 1:Tmax, x= x)
          },
          .id="rep")
}
models[names(models) != true_model]

true_sim <- sim(models[[true_model]], 5, 100, 1) %>%
  mutate(model = true_model, mean = x, ymin = NA, ymax = NA)
df <- map_dfr(models[1:2], sim, 5, 100, 100, .id = "model")
forecast <- df %>% 
  group_by(model, t) %>% summarise(mean = mean(x), sd = sd(x))

forecast %>%
  ggplot(aes(t, mean, col=model)) +
  geom_line() + 
  geom_line(data = true_sim) + 
  geom_ribbon(aes(fill= model, ymin = mean - 2*sd, ymax = mean + 2*sd), alpha=0.1, col=NA)

```

```{r}
d <- forecast %>% group_by(model) %>% 
  mutate(true = true_sim$x) %>%
  summarise(rmsd = sqrt(mean( (mean - true) ^ 2 )),
            r2 = 1 - sum( (mean - true) ^ 2  ) / sum( mean ^ 2 )
            )
rmsd <- d %>% pull(rmsd)
names(rmsd) <- d %>% pull(model)

opt <- sqrt(mean((true_sim$x - 10)^2))
opt
```

Model 2 clearly outperforms model 1 in predictive capacity. The root mean squared deviation between the model 2 mean and the realization is `r rmsd[["f2"]]`, far lower than model 1 at `r rmsd[["f1"]]`.

```{r}
# Simulate the evolution of probability distribution directly using a matrix exponent
prob_dynamics <- function(M, X, Tmax){
  probability <- t(M) %^% Tmax %*% X 
  data.frame(state = states, probability)
}

x0 <- which.min(abs(states - 6))
X <- numeric(length=length(states))
X[x0] <- 1

map_dfr(transitions, 
        function(m) prob_dynamics(m[,,1], X, 100),
        .id = "model") %>%
  group_by(model) %>%
  ggplot(aes(state, probability, col=model)) +
  geom_line()
```