### 2021-03-31 manuscript/appendix.Rmd
consolidate appendix

---
title: 'Appendix A'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

---

## Optimal quotas as a Markov Decision Process (MDP) problem

The question of optimal harvest quotas given a population growth model can be posed as a Markov Decision Process (MDP). @Marescot2013 provides an accessible overview of MDP problems and their solutions in conservation context.  @Reed1979 provides a formal proof of optimal harvest strategy for stochastic dynamics under the assumption of any concave growth function. The problem can be summarized as follows:

A manager seeks to maximize the sum of the utility derived from such a harvest
and such a state, $U(X_t,H_t)$, over all time, subject to discount rate $\delta$:

\begin{equation}
\sum_{t=0}^{t=\infty} U(X_t, H_t) \delta^t
\end{equation}

While in principle the utility could reflect many things, including the cost of fishing, market responses to supply and demand, the value of recreational fishing, the intrinsic value fish left in the sea [see @Halpern2013], for simplicity we will assume utility is merely a linear function of the harvest quota set by the manager, i.e. a fixed price $p$ per kilogram of fish harvested: $U(X_t, H_t) = p \min(H_t, X_t)$ (noting that realized harvest cannot exceed the available stock). As units are already specified in dimensionless quantities in this example, so without loss of generality we will set $p=1$.

To solve this optimization, we must also know the dynamics of $X_t$, describing how future stock $X_{t+1}$ changes from the present state, $X_t$, after to harvest $H_t$:

\begin{equation}
X_{t+1} = f(X_t, H_t)
\end{equation}

This problem is already well studied, and it is worth noting that even under the simple utility function which assigns no value to fish that are not harvested, the optimal strategy still seeks to sustain the fish population indefinitely; as @Clark1973 shows for the deterministic function $f$ and @Reed1979 extended to the stochastic case.  As @Reed1979 proves, the optimal strategy can be characterized by a policy of "constant escapement," in which the manager adjusts the harvest effort each year in an effort to ensure the same number or biomass of fish are left in the sea each year.  The optimization problem can then be solved by dynamic programming, using the recursive Bellman equation. This technique, known as stochastic dynamic programming (SDP) is widely used in conservation decision problems [e.g. @Mangel1985, @Marescot2013].

It is worth noting that while constant escapement is used in the management of several important fish stocks such as salmon, most marine fisheries are not managed using this optimal solution for a fluctuating population, but instead rely on a 'constant harvest effort' (or constant "yield") policy.  In the case of a deterministic model at equilibrium, these policies are identical.  Though constant-effort policies are technically sub-optimal for the problem as stated above with stochastic growth, in practice they can be more robust, such as when measurements of the current stock size $X_t$ are uncertain.  These issues are discussed in detail in @pomdp-intro.  For our purposes, it suffices to note that constant escapement policy is relevant both from a theoretical standpoint as the optimal solution for the problem under consideration, and an applied standpoint as a policy that guides management of many salmon fisheries.  The optimal control approach using SDP as illustrated here is used in a wide and growing number of conservation problems.

Bear in mind that in the example presented here, SDP merely serves the role of a convenient and established way for a manager to determine the optimal action, given some model estimation.  Whether we use SDP or some other way to determine the optimal policy, given the model, is immaterial to the conclusion.

## Solving MDP via SDP

To solve the decision problem using SDP, we define the state space and the action space on a discrete grid of 240 points.  The maximum state is set well above the largest carrying capacity used in the models, which limits the influence of boundary effects introduced by the transformation to a discrete, finite grid.  Available harvest actions match the state space, effectively allowing any harvest level to be possible.

```{r model_definitions}
states <- seq(0,24, length.out = 240)
actions <- states
obs <- states
```

The utility (reward) of an action is set to the (realized) harvest (e.g.
a fixed price per unit fish, with no cost applied to harvest effort).  Future
utility is discounted by fixed factor of $0.99$. Classic maximum sustainable 
yield models [@Schaefer1954] ignore discounting, while modern economic optimization
models insist on it, so a small discount conforms to the latter while reasonably
approximating the former.  The qualitative conclusion is not sensitive to the 
discounting rate.

```{r}
reward_fn <- function(x,h) pmin(x,h)
discount <- 0.99
```

Alternate reward functions with varying price and cost structures are possible but do not qualitatively impact the conclusions.  This reward function is a limiting case of any more complex reward, and corresponds both to classic work that does not model utility explicitly [e.g. MSY theory, @Schaefer1954], as well as explicit assumptions typically made in more recent models [@Reed1979]. Moreover, using a simple reward function makes it clear that model that performs best is not doing so merely because of particular features baked into the a carefully chosen reward rule.

# Ecological Models

The manager chooses between two different logistic growth models, each of which
can be thought of as an approximation to the underlying "true" population model:

\begin{equation}
f_i(Y) = Y + Y r_i \left( 1 - \frac{Y}{K_i} \right) \xi_i(t)
\end{equation}

where $\xi_i(t)$ is a multiplicative log-normal noise term with mean 1 and log-standar-deviation $\sigma_i$.

Model 1 has $r_1 = 2$, $K = 16$, and $\sigma_1 = 0.05$.
Model 2 has $r_2 = 0.5, K_2 = 10$, and $\sigma_2 = 0.075$.

Meanwhile, the true population growth rate is simulated using a function with
non-linear per-capita growth:

\begin{equation}
f_i(Y) = Y + Y^4 r_i \left( 1 - \frac{Y}{K_i} \right)
\end{equation}

with $r_3 = 0.002$, $K_3 = 10$ and $\sigma_3 = 0.05$.

Because the methods applied here: Stochastic Dynamic Programming, Iterative Forecasts assessed with Proper Scoring Rules, or Adaptive Management, are not themselves novel, the reader is referred to previous reviews and text books which can provide a thorough introduction to methods, in particular, @Marescot2013 and @Smith1981.  However, implementation of these approaches generally depends on numerical methods, for which even precise mathematical formulae or pseudo-code descriptions can be inadequate to specify unambiguously or to properly facilitate replication [@Barnes2010].  As such, this appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulae presented here, it is also less ambiguous. Though R is already widely used among ecologists, the implementations concern mostly matrix algebra that can be translated to other languages, and the `MDPtoolbox` package used here is also implemented in Matlab [@Marescot2013; @MDPtoolbox].

Below, I provide annotated code necessary to completely reproduce all of the analysis presented in the main paper. This analysis is run in R [@R] uses `MDPtoolbox` [@MDPtoolbox] for solving Markov Decision Processes (MDP) using stochastic dynamic programming functionality, `expm` for matrix exponentials [@expm], and a few custom MDP functions provided by our package, `mdplearning` [@mdplearning].  We will also use `tidyverse`  packages for basic manipulation and plotting [@tidyverse]. This file is also available as an RMarkdown document [@rmarkdown] at  <https://github.com/cboettig/bad-forecast-good-decision>.  The code below generates the data tables required for each plot shown in the main paper.  Data are stored as `.csv` tables to be accessible to any relevant software program. Plotting commands used in the main paper are embedded in the paper's RMarkdown file at <https://github.com/cboettig/bad-forecast-good-decision>.

```{r graphics_setup, message=FALSE, warning=FALSE, include = FALSE}
## Plotting themes, colors, and fonts 
## aesthetics choices only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
library(styler)
extrafont::loadfonts(quiet = TRUE)
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(8)
txtcolor <- "#586e75"

knitr::opts_chunk$set(cache=FALSE, tidy = "styler", message = FALSE, warning = FALSE, echo = TRUE)

```

```{r setup, message=FALSE}
library(tidyverse)
library(MDPtoolbox)
library(expm)
# remotes::install_github("boettiger-lab/mdplearning")
library(mdplearning)
```

# Model definitions

```{r}
# K is at twice max of f3; 8 * K_3 / 5
f1 <- function(x, h = 0, r = 2, K = 10 * 8 / 5){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}
f2 <- function(x, h = 0, r = 0.5, K = 10){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}

# max is at 4 * K / 5 
f3  <- function(x, h = 0, r = .002, K = 10){
  s <- pmax(x - h, 0)
  s + s ^ 4 * r * (1 - s / K)
}

## gather models together, indicate true model
sigma_g <- 0.05
models <- list("1" = f1, "2" = f2, "3" = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)
true_model <- "3"
```

On a discrete grid of possible states and actions, we can define the growth rate of a given state $X_t$ subject to harvest $H_t$, $f(X_t,H_t)$ as set of matrices.
Each matrix $i$ gives the transition probabilities for any current state to any future state, given that action $i$ is taken.

```{r transition_matrices}
transition_matrices <- function(f, states, actions, sigma_g){
  n_s <- length(states)
  n_a <- length(actions)
  transition <- array(0, dim = c(n_s, n_s, n_a))
  for (k in 1:n_s) {
    for (i in 1:n_a) {
      nextpop <- f(states[k], actions[i])
      if(nextpop <= 0){
        transition[k, , i] <- c(1, rep(0, n_s - 1))
      } else if(sigma_g > 0){
        x <- dlnorm(states, log(nextpop), sdlog = sigma_g)
        if(sum(x) == 0){ ## nextpop is computationally zero
          transition[k, , i] <- c(1, rep(0, n_s - 1))
        } else {
          x <- x / sum(x) # normalize evenly
          transition[k, , i] <- x
        }
      }
    }
  }
  transition
}
```

This follows the standard setup for standard stochastic dynamic programming, see @Marescot2013. Having defined a function to compute the transition matrix, we can use it to create matrices corresponding to each of the three models:

```{r}
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("1", "2", "3")

```

Likewise, a corresponding matrix defining the rewards associated with each state $X$ and each harvest action $H$ can also be defined.

````{r}
## Compute reward matrix (shared across all models)
n_s <- length(states)
n_a <- length(actions)
reward <- array(0, dim = c(n_s, n_a))
for (k in 1:n_s) {
  for (i in 1:n_a) {
    reward[k, i] <- reward_fn(states[k], actions[i])
  }
}
```

# Optimal control solutions

We use value iteration to solve the stochastic dynamic program [@Marescot2013; @MDPtoolbox] for each model. This determines the optimal harvest policy for each possible state, given each model. Because this step is the most computationally intensive routine, we cache the results using memosization conditioned on the transition matrices [@memoise].  Running this code with alternate transition matrices automatically invalidates that cache, reducing the risk of loading spurious results.

```{r sdp, results="hide"}
mdp <- memoise::memoise(mdp_value_iteration,
                        cache = memoise::cache_filesystem("cache/"))

policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp(P, reward, discount = discount, 
                epsilon = 0.01, max_iter = 2000, V0 = rep(0,dim(P)[[1]]))
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")

write_csv(policies, "../data/policies.csv")
```

# Simulations and step-ahead forecasts

We simulate fishing dynamics under the optimal policy for each model, using a simple helper function from the `mdplearning` package. Because growth dynamics are stochastic, we perform 100 simulations of each model from identical starting condition to ensure results are not the result of chance alone.

```{r simulations}
library(mdplearning)
Tmax <- 100
x0 <- which.min(abs(states - 6))
reps <- 100
set.seed(12345)

## Simulate each policy reps times, with `3` as the true model:
simulate_policy <- function(i, policy){
  mdp_planning(transitions[[true_model]], reward, discount,
           policy = policy, x0 = x0, Tmax = Tmax) %>%
    select(value, state_index = state, time, action_index = action)  %>% 
    mutate(state = states[state_index])
}

sims <- 
  map_dfr(names(transitions), 
          function(m){
            policy <- policies %>% filter(model == m) %>% pull(policy)
            map_dfr(1:reps, simulate_policy, policy = policy, .id = "reps")
          },
          .id = "model"
         )

write_csv(sims, "../data/sims.csv")
```

Using the transition matrices directly, we can examine what each model would 
have forecast the future stock size to be in the following year when no fishing
occurs (note that for each model, we use the transition matrix that corresponds
to 'no fishing', `model[[state_index, ,1]]`) (Fig 1a, main text).

The transition matrices give the full (discretized) probability distribution,
from which we can easily calculate both the expected value and the 95% 
confidence interval.

```{r stepahead_unfished}
stepahead_unfished <- sims
stepahead_unfished$state_index <- rep(sims$state_index[sims$model == "1"],3)

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ]) 
```

We also look at the forecast each model makes when implementing the corresponding
optimal harvest:

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state]) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

# Proper scores

It is straight forward to apply the proper scoring formula of @Gneiting2007 
based on the first two moments of the distribution to score the respective
forecasts under both the unfished and actively managed scenario sfor each model:

```{r proper_scores, message = FALSE}
# Gneiting & Raferty (2007), eq27
scoring_fn <- function(x, mu, sigma){ -(mu - x )^2 / sigma^2  - log(sigma)}

stepahead_unfished <- stepahead_unfished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))

stepahead_fished <- stepahead_fished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))
```

We store the replicate predictions under each model at each timestep, along with the scores, in a `predictions.csv` table for plotting.

```{r}
predictions <- 
  stepahead_unfished %>% 
  select(time, model, reps, expected, low, high, true, score) %>% 
  mutate(scenario = "A_unfished") %>% 
  bind_rows(stepahead_fished  %>% 
          select(time, model, reps, expected, low, high, true, score) %>% 
          mutate(scenario = "B_fished")
  ) %>% 
  mutate(model = as.character(model))

write_csv(predictions, "../data/predictions.csv")
```

# Adaptive Management

Passive adaptive management using a Bayesian learning scheme still learns the wrong model. Adaptive management for MDP problems is particularly computationally intensive, and not directly supported by the `MDPtoolbox` implementation, though the principle involves only a relatively straightforward, if not particularly efficient extension to the task of solving and MDP problem for a single model shown above.  Given a set of transition probability matrices $P_i$ and a belief probability $b_i$ that the $i$th model is correct, it simple to define the transition matrices over possible models by the laws of conditional probability, $P = \sum_i P_i b_i$, given $\sum_i b_i = 1$. After each subsequent action and then new observation x_j, the $b_i$ probabilities for each model are updated according to Bayes law: if $b_i = P(M_i)$ is the prior probability that the model $M_i$ is correct, then the posterior probability that model $M_i$ is correct given observation $x_j$ is:

$$b_i' = P(M_i | x_j) = P(M_i) P(x_j | M_i) = b_i P(x_j | M_i) $$

Note that because the transition matrices are now altered after every action and observation, it is necessary to repeat the complete SDP calculation to determine the optimal action given the updated uncertainty.  While straight forward, this is what makes adaptive management for MDP problems particularly computationally intensive and thus poorly suited for handling a large space of candidate models. Nevertheless, our example of two models is most feasible, and even the larger example of 36 models is not unreasonable. Here I make use of the `mdplearning` R package, which provides a straight-forward implementation of this algorithm.

```{r}
adaptive_management <- memoise::memoise(mdp_learning, cache = memoise::cache_filesystem("cache/"))
am1 <- adaptive_management(transitions[1:2], reward, discount, 
                           model_prior = c(0.99, 0.01), x0 = x0, 
                           Tmax = 50, true_transition = transitions[[3]], 
                            epsilon = 0.001, max_iter = 2000)
```

I also compare the adaptive management solution to an approach which still integrates uncertainty over both models, but treats the resulting policy as fixed for the duration of management, rather than adaptive _learning_ by updating probabilities.  This approach is sometimes referred to as uncertainty "planning," as it accounts for the same initial uncertainty over models but does not attempt to learn.   For consistency, I use the same model prior as in the learning case, which heavily favors model 1 to begin with.  Not surprisingly, the resulting policy is nearly identical to that under model 1 alone.

```{r}
# Compute policy using SDP over both models, given fixed prior beliefs
policy_planning <- memoise::memoise(mdp_compute_policy, cache = memoise::cache_filesystem("cache/"))
policy <- policy_planning(transitions[1:2],  
                          reward = reward, 
                          discount = discount,
                          model_prior = c(0.99, 0.01),
                          Tmax = 50, 
                          epsilon = 0.001, 
                          max_iter = 2000)
```

```{r}
# Simulate under true model using fixed planning policy
non_am <- mdp_planning(transitions[[3]],  
                       reward = reward, 
                       discount = discount,
                       model_prior = c(0.99, 0.01),
                       x0 = x0, 
                       a0 = 1,
                       policy = policy$policy,
                       Tmax = 50)  %>% 
  mutate(belief = 0.99, # belief in model 1 is fixed
         method = "planning")
```

We combine results of the two methods and store the resulting `data.frame` of stock size, quota, and belief in Model 1 over time as `am.csv` for later plotting.

```{r}
am <- am1$df %>%
  mutate(belief = am1$posterior$V1,
         method = "learning") %>%
  bind_rows(non_am) %>%
  mutate(stock = states[state],
         quota = actions[action]) %>%
  select(time, stock, quota, belief, method)

write_csv(am, "../data/am.csv")

```

## Adaptive management over many models

We repeat the analysis above over a grid of $r$ and $K$ values.

```{r}
# compare over grid set of candidate r and K
rs <- c(0.01, 0.1, 0.2, 0.3, 0.4, 0.5, 2)
Ks <- c(0.5, 1, 1.5, 4, 10, 16)
sigma <- 0.1
n_models <- length(rs)*length(Ks)
pars <- expand.grid(rs, Ks)
names(pars) <- c("r", "K")
pars$sigma <- 0.1
closure <- function(r, K){
  function(x, h = 0){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
  }
}
many_models <-  lapply(1:n_models, function(i) closure(pars[i,"r"], pars[i,"K"]))

# transition matrices
many_transitions <- lapply(seq_along(many_models), function(i) 
  transition_matrices(many_models[[i]], states,  actions, sigma))
names(many_transitions) <- as.character(seq_along(many_models))
true_transition <- transition_matrices(f3, states, actions, sigma_g)

## Once again, initial belief favors "model 1"
model_prior <- c(rep(.01, n_models-1) / (n_models - 1), .99)

```

With the transition matrices defined for all 42 models, we perform adaptive management as before:

```{r}
adaptive_management <- memoise::memoise(mdp_learning, cache = memoise::cache_filesystem("cache/"))
am_many <- adaptive_management(many_transitions,  
                               reward = reward, 
                               discount = discount,
                               model_prior = model_prior,
                               x0 = x0, 
                               Tmax = 20, 
                               true_transition = true_transition, 
                               epsilon = 0.001, 
                               max_iter = 2000)

```

As before, we can compare the adaptive learning case to the policy which accounts for the initial uncertainty over all `r n_models` models, but does not seek to learn.

```{r}
policy_many <- policy_planning(many_transitions, reward = reward,  discount = discount,
                          model_prior = model_prior, Tmax = 20, 
                          epsilon = 0.001, max_iter = 2000)

non_am_many <- mdp_planning(true_transition,  
                       reward = reward, 
                       discount = discount,
                       model_prior = model_prior,
                       x0 = x0, 
                       a0 = 1,
                       policy = policy_many$policy,
                       Tmax = 20) %>% 
  mutate(belief = 0.99, # belief in model 1 is fixed
         method = "planning")

model1 <- dim(am_many$posterior)[2]

am_multi <- am_many$df %>% 
  mutate(belief = am_many$posterior[,model1],
         method = "learning") %>%
  bind_rows(non_am_many) %>%
  mutate(stock = states[state],
         quota = actions[action]) %>%
  select(time, belief, stock, quota, method)

write_csv(am_multi, "../data/am_multi.csv")
```

# Posteriors

```{r}
posteriors <- data.frame(r = rs, K = Ks, t(am_many$posterior)) %>%
  pivot_longer(starts_with("X"), 
               values_to = "probability",
               names_to = "time") %>% 
  mutate(time = as.integer(as.factor(time)))

i <- 20
r_marginals <- posteriors %>% group_by(time,r) %>% 
  summarise(prob = sum(probability), .groups = "drop") %>% 
  filter(time %in% i)

r_marginals %>%  ggplot(aes(r, prob, group=time)) + geom_bar(stat="identity") + facet_wrap(~time)
```

## Marginal Posterior Probability evolution over K

```{r}
i = 20
K_marginals <- posteriors %>% group_by(time,K) %>% 
  summarise(prob = sum(probability), .groups = "drop") %>% 
  filter(time %in% i) 
K_marginals %>%
  ggplot(aes(K, prob, group=time)) + geom_bar(stat="identity") + facet_wrap(~time) + scale_x_log10()

```

# Economic value tabulation

To plot the economic value over time, we must sum up the discounted values
at each time step, and then average over replicate simulations of each model:

```{r plot_npv}
##  Net Present Value accumulates over time
npv_df <- sims %>% 
  group_by(model, reps) %>%
  mutate(npv = cumsum(value * discount ^ time)) %>%
  group_by(time, model)  %>% 
  summarise(mean_npv = mean(npv), .groups="drop") %>% 
  arrange(model, time)

write_csv(npv_df, "../data/npv_df.csv")
```

```{r}
# tabular comparisons
npv <- npv_df %>% group_by(model) %>% summarize(npv = max(mean_npv))

am_npv <- sum(am$value * discount ^ am$time)
am_economics_percent <- round(am_npv / npv[[1,"npv"]] * 100)

mean_state <- sims %>% group_by(model) %>% summarize(state = mean(state))
am_ecology_percent <- round(mean(states[am$state])/mean_state[[1, "state"]]*100)
```

Under adaptive management, the manager realizes only `r am_economics_percent`%
of the economic value that would be achieved under Model 1 alone, and only 
`r am_ecology_percent`% of the spawning stock biomass that would have been 
achieved under Model 1 alone.

# Growth models

To plot growth curves of individual models (Fig 4a), we evaluate

$$\Delta x = x_{t+1} - x_t = f(x_t) - x_t$$

for each model for all possible states $x_t$.

```{r plot_models}
model_curves <- 
  map_dfc(models, function(f) f(states) - states) %>%
  mutate(state = states) %>%  pivot_longer(names(models), "model")

write_csv(model_curves, "../data/model_curves.csv")

```

## Growth models in the many-models case:

```{r plot_models, fig.width=6, fig.height=4}
final_belief <- t(am_many$posterior)
final_belief <- final_belief[,dim(final_belief)[[2]]]
posteriors <- data.frame(model = as.character(1:length(many_models)), belief = final_belief)

model_set <-c(f3,  many_models)
names(model_set) <- c("true", 1:length(many_models))
d <- 
  map_dfc(model_set, function(f) f(states) - states) %>%
  mutate(state = states) %>% pivot_longer(names(model_set[-1]), "model") %>%
  left_join(posteriors)

d %>% 
  ggplot(aes(state, value, group=model, alpha = belief)) +
  #geom_hline(aes(yintercept = 0), lwd=1, lty=2) + 
  geom_line(lwd=1.5) + 
  geom_line(aes(state, true, alpha=1), col = "red", show.legend = FALSE) +
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x") 
```

# References