### 2021-04-13 reviews/README.Rmd
revisions

---
output: github_document
---

Revision reply letters are encrypted on this repository.  The reply letter quotes the 
recommendations of the editors and reviewers, to which those individuals hold copyright
unless they grant express permission to reproduce their comments here.  Encrypting
the reply letter preserves this confidential correspondence in this otherwise-public repository.
The source `.Rmd` and resulting `.pdf` reply letter are then `gitignore`'d.

A template .Rmd of the reply letter, capturing my formatting settings, is included here.

Encrypt to my public key:

```{r setup}
library(gpg)
gpg::gpg_recv("3908E1CFD28B380C") # import, needed only once
```

```{r}
msg <- gpg_encrypt("review-reply.Rmd", receiver =  "3908E1CFD28B380C")
writeLines(msg, "review-reply.Rmd.gpg")
```

## decrypt

(decrypt locally offline where my private key is available from yubikey, `gpg -d review-reply.Rmd.gpg`)