### 2021-05-18 notebook/model-estimates.Rmd

---
title: "Data Sources"
output: github_document
---

```{r}
library(contentid)
library(tidyverse)
library(httr)
#library(taxalight)
```

# May model fit

# Logistic model fit

# Allen model fit

# Transition matrices from ARIMA?