### 2021-05-19 notebook/model-estimates.Rmd
:construction:

---
title: "Data Sources"
output: github_document
---

```{r}
library(contentid)
library(tidyverse)
```

# May model fit

---
title: 'Appendix B'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

---

```{r graphics_setup, message=FALSE, warning=FALSE, include = FALSE}
## Plotting themes, colors, and fonts 
## aesthetics choices only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
library(styler)
extrafont::loadfonts(quiet = TRUE)
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(8)
txtcolor <- "#586e75"

knitr::opts_chunk$set(cache=FALSE, tidy = "styler", message = FALSE, warning = FALSE, echo = FALSE)

```

```{r setup, message=FALSE}
library(tidyverse)
library(nimble)
```

```{r}
siteA <- read_csv("barbary_macaque.csv") %>% 
  select(SampleYear, PopulationUntransformed, LocationID) %>%
  mutate(LocationID = as.character(LocationID)) %>% 
  filter(LocationID == "414")

scale <- 20
x <- siteA$PopulationUntransformed / scale
constants <- list(N = length(siteA$SampleYear),
                  x0 = x[[1]]
)
```

```{r}
logistic  <- nimble::nimbleCode({
  r ~ dunif(0,2)
  K ~ dunif(0,10)
  sigma ~ dunif(0,.4)
  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K)
    x[t+1] ~ dlnorm(log(mu[t]), sd = sigma)
  }

})

```

```{r}
may  <- nimble::nimbleCode({
  x[1] <- x0
  q <- 3
  s <- 2
  r ~ dlnorm(log(.7) - s^2/2, s)
  K ~ dlnorm(log(1.2)- s^2/2, s)
  a ~ dlnorm(log(0.2) - s^2/2,  s)
  b ~ dlnorm(log(0.15) - s^2/2, s)
  sigma ~ dlnorm(log(0.05) - s^2/2, s)

  for(t in 1:(N-1)){
    mu[t] <-  x[t] + x[t] * r * (1 - x[t] / K)  - a * x[t] ^ q / (x[t] ^ q + b ^ q)
    x[t+1] ~ dnorm(log( mu[t] ), sd = sigma)
  }

})

```

```{r}
mcmc <- function(model, x, constants, niter = 100000, nburin = 1000){
  r_model <- nimbleModel(model, constants = constants)
  r_model$setData(list(x = x))
  c_model <- compileNimble(r_model)
  r_mcmc <- buildMCMC(c_model)
  c_mcmc <- compileNimble(r_mcmc)
  df <- runMCMC(c_mcmc, niter = niter, nburnin = nburin)
  tibble::as_tibble(df)
}
```

```{r}
l_draws <- mcmc(logistic, x, constants)
l_draws %>% summarise(across(everything(), list(mean = mean, sd = sd)))

```
```{r}
bayesplot::mcmc_trace(l_draws)
```

```{r}
m_draws <- mcmc(may, x, constants)
m_draws %>% summarise(across(everything(), list(mean = mean, sd = sd)))
bayesplot::mcmc_trace(m_draws)

```

# Logistic model fit

# Allen model fit

# Transition matrices from ARIMA?