### 2021-08-23 manuscript/manuscript.Rmd
[WIP] :pen:

---
title: 'The forecast trap'
- name: Carl Boettiger
  email: cboettig@berkeley.edu
  affiliation: a,1
date: "`r Sys.Date()`"

abstract: |
  Encouraged by decision makers' appetite for future information on topics ranging from elections to pandemics, and enabled by the explosion of data and computational methods, model based forecasts have garnered increasing influence on a breadth of decisions in modern society. Using a classic example from fisheries management, I demonstrate that selecting the model or models that produce the most accurate and precise forecast (as measured by most widely used statistical scores) can lead to decidedly worse outcomes (as measured by real-world objectives such as dollars or fish). This can create a forecast trap: in which the manager not only over-exploits the fishery leading to declining biomass and declining yields, but also becomes increasingly convinced that these actions are not over-exploitation but in fact consistent with the best models and data for sustainable management. The forecast trap is not unique to this example, but possible whenever (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not itself in our candidate set of models.

corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: cboettig@berkeley.edu"

bibliography: refs.bib
address:
- code: a
  address: Department of Environmental Science, Policy, and Management, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA

keywords:
- forecasting,
- adaptive management,
- stochasticity,
- uncertainty,
- optimal control

header-includes:
  - \linenumbers
  - \usepackage{endfloat}

csl: csl/ecology-letters.csl
output: rticles::elsevier_article
layout: 3p
journal: Ecology Letters
---

- abbreviated running title: "The forecast trap"
- Authorship statement: _This is a single-author paper_
- data accessibility statement: _All simulation data generated for analyses here, along with code required for the analysis, is available in the Zenodo Data archive, https://doi.org/10.5281/zenodo.4660621_
- number of words in the abstract: 116
- number of words in main text: 4998
- number of cited references: 33
- number of tables & figures: 4 Figures

\pagebreak

```{r, message=FALSE, warning=FALSE, include = FALSE}
## Plotting themes, colors, and fonts 
## aesthetics choices only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
library(styler)
extrafont::loadfonts(quiet = TRUE)
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(8)
txtcolor <- "#586e75"
```

```{r, include=FALSE}
rmarkdown::render("appendix.Rmd")
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, 
                      warning = FALSE, dev="cairo_pdf", fig.height=5, fig.width=7.5)
```

Global change issues are complex and outcomes are difficult to predict [@Clark2001].
To guide decisions in an uncertain world, researchers and decision makers may consider a range of alternative plausible models to better reflect what we do and do not know about the processes involved [@Polasky2011].
Forecasts or predictions from possible models can indicate what outcomes are most likely to result under what decisions or actions.
This has made model-based forecasts a cornerstone for scientifically based decision making.
By comparing outcomes predicted by a model to future observations, a decision maker can not only _plan_ for the uncertainty, but also _learn_ which models are most trustworthy. 
The value of iterative learning has long been reflected in the theory of adaptive management [@Walters1978] as well as in actual adaptive management practices such as Management Strategy Evaluation (MSE) [@Punt2016] used in fisheries, and is a central tenet of a rapidly growing interest in ecological forecasting [@Dietze2018].
But, do iterative learning approaches always lead to better decisions?

In this paper, I demonstrate that the model that makes the better prediction (rigorously defined as a strictly proper score, @Gneiting2007) is not necessarily the model that makes the better policy (rigorously defined in terms of utility, e.g. expected net present value, @Clark1990).
I show that our best methods for learning about model structure or parameters by repeatedly comparing forecasts to observations can be counter-productive.
Put another way, the value of information (VOI, as measured by the expected utility given that information minus the utility without it; see @Howard1966; @Katz1987), can actually be negative.
When VOI is negative, the decision-maker may become trapped into accepting mediocre outcomes derived from a model that makes accurate forecasts, even when a less accurate model that would generate better outcomes is available.
In our example, the manager will decide the fishery in question simply has low productivity, because such a model yields better predictions, rather than realizing that the low productivity observed is in fact a consequence of the over-harvesting.
This disconcerting situation can arise whenever two conditions are met: (1) the optimal management policy is not unique to the generative process, and (2) the generative process is not included in candidate set of models.
These conditions do not guarantee the trap will occur, only the circumstances in which it cannot be ruled out entirely.

The forecast trap is not the only mechanism by which some model-choice methods lead to worse outcomes.
Previous work has long acknowledged the panoply of ways in which model-based decision making can go astray due to conflicting incentives, implementation errors, or lack of resources for monitoring and updating [e.g. @Ludwig1993].
Another widely recognized problem is that of over-fitting [@Burnham1998], in which the model that best fits historical data fails to best predict future data [@Ginzburg2004]. 
Under such circumstances, it is easy to see how an over-fit model would also lead to bad outcomes.
However, over-fitting plays no role in the forecast trap, where model predictions are assessed only using probabilistic forecasts, and not observations which had previously been used to fit the models. 
Formally, these scores satisfy the 'proper scoring' rule of @Gneiting2007, which proves no other probabilistic prediction $Q(x)$ will have a better expected score than that of the generative process $P(x)$. 
@Gneiting2007's proof of proper scoring has since become a critical tool to avoid over-fitting when choosing models to make decisions, but as I illustrate, will not prevent the forecast trap.

# From Predictive Models to Decision Policies

How do we translate a model-based forecast into a decision?
It is impossible to discuss outcomes associated with a forecast without first agreeing on this process.
In practice, decision-makers may use a forecast in a wide variety of ways in selecting a course of action, including ways which may run counter to the stated objectives of management [@Ludwig1993].
In principle at least, the field of decision theory provides a formal mechanism for determining the optimal strategy given a model forecast.
For instance, a wide range of ecological conservation and management problems can be expressed as a Markov Decision Process (MDP) problems [@Marescot2013].
Existing computer algorithms such as stochastic dynamic programming (SDP) take a probabilistic model _forecast_ (more precisely, the probability $P(x_{t+1} | x_t, a_t)$ of the system being in state $x_{t+1}$ in the next iteration given that it was previously in state $x_t$ and the manager selected action $a_t$) and the _desired management objective_ (i.e. the maximize the expected biomass of species protected or the expected dollar profit of a fishery [see @Clark1990; @Halpern2013]) as input, and return the _decision policy_ which maximizes that objective [@Marescot2013].
This provides a principled way to associate a decision policy with any given forecast model.

Two features of this approach are worth emphasizing. 
First, the resulting decision is derived directly from the forecast model and the desired objective.
The SDP algorithm is a reasonable description of the approach any ideal manager would use -- considering all possible outcomes from all possible sequences of actions and selecting the best sequence.
For complex models this process is too laborious even for a computer, and is often simplified by considering only a selection of predetermined policies (as in Management Strategy Evaluation, MSE, @Punt2016), or scenarios (as in scenario analysis, @Polasky2011).
Such shortcuts are often necessary for complex real-world models, but open additional room for error: the policy we derive from a given forecast may perform poorly not because the model forecast was at fault, but because of those simplifying assumptions about possible policies. 
To ensure that the forecast trap is not a result of such assumptions about possible policies, we will consider a problem simple enough to solve directly with SDP. 
This leads to the second point: the resulting decision policy is optimal, so long as the forecast model is correct.
In this way, the SDP merely stands in for a mathematically precise way in which forecasts are turned into decisions.
Recognizing that the SDP-derived policy (A) comes directly from the forecast model, and (B) gives the optimal policy for said forecast, seems to suggest that the whatever model makes the better forecast will surely also lead to better outcomes (as measured in terms of whatever utility we have chosen to maximize). 
While this intuition is no doubt _often_ accurate, our purpose here is to demonstrate that it is by no means _guaranteed_:
it is also possible for the model which makes the better forecast to lead to worse outcomes.

# Ecological Models

I illustrate this problem using an example from fisheries management. 
Fisheries are a significant economic and conservation concern worldwide and their management remains an important debate [e.g. @Worm2006; @Worm2009; @Costello2016].
Moreover, their management has been a proving grounds for theoretical and practical decision-making issues which are widely applicable in other areas of ecology and conservation [@Ludwig1993; @Lande1994], and one that has long wrestled with issues of uncertainty in the context of management decisions [e.g. @Clark1973; @Reed1979; @Walters1981; @Ludwig1982].

While methods such as iterative forecasting [@Deitze2018] and adaptive management [@Walters1978] can be _applied_ to real-world using empirical data, we can only _evaluate_ their potential in hypothetical examples when the true model is known, e.g. through numerical simulation.
That approach allows us to compare both predictions and outcomes across implementations in independent identical replicate worlds.
As noted above, we will assume our underlying model simple enough to solve by SDP, ensuring any poor outcomes from a given forecast are not merely an artifact of an imperfect decision process.
Simple models also have the virtue in being accessible to closed form analysis, which, as we shall see, can give greater insight into when and why this forecast trap arises.
That insight will in turn will allow us to examine if the same problem is likely to arise under more complex models.

<!--
While modern fisheries management frequently relies on complex models which may contain over 100 parameters to reflect the specific age or stage structure of a specific fish stock [@ramlegacy2018; @ram], fisheries management and research has always appreciated the central insights offered by simpler one-dimensional models which continue to underpin both the theory and practice.
For example, the strategies of constant mortality [@Schaefer1954] and constant escapement [@Clark1973; @Reed1979] are built on the same class of models considered here, which also continue to underpin global analyses of stock-rebuilding [e.g. @Costello2016; @Memarzadeh2019] and the foundations of adaptive management [e.g. @Walters1978; @Smith1981; @Walters1981].
While more complex models are often required for specific applications, such models both share features with and are guided by theory built on much more general, simpler models [@Levins1966],
from the Maximum Sustainable Yield of @Schaefer1954 or the epidemiological reproductive number, $R_0$ of @Kermack1927. 
-->

The sustainable harvest decision problem can be stated as follows: The fish stock is observed to be in state $X_t$ at time $t$, and is then subjected to some harvest $H_t$ before recruiting new fish, subject to stochastic environmental noise $\xi_t$, to bring the stock to $X_t+1$,

\begin{equation}
X_{t+1} = f(X_t - H_t, \xi_t) 
\end{equation}

A manager seeks each year to select the harvest quota $H_t$ which will maximize the sum
of the utility derived from such a harvest and such a state, $U(X_t,H_t)$, over all time,
subject to discount rate $\delta$ [@Clark1973]:

\begin{equation}
\sum_{t=0}^{t=\infty} U(X_t, H_t) \delta^t \label{utility}
\end{equation}

We will assume we have been given a fixed price of fish $p=1$ with no additional cost on additional harvest, $U(X_t, H_t) = p \min(H_t, X_t)$ modest discount $\delta = 0.99$.

Let us assume that our set of candidate models are simply the possible parameterizations of a stochastic version of the classic Gordon-Schaefer model [@Schaefer1954; @Gordon1954]:

\begin{equation}
f_i(Y_{t+1}) = Y_t + r_i Y_t \left( 1 - \frac{Y_t}{K_i} \right) * \xi_t(\sigma)
\end{equation}

Where $Y$ is the population size after harvest, $Y_t = X_t-H_t$  and $\xi_t(\sigma)$ represents log-normal random noise with a mean of unity and log-standard-deviation $\sigma_i$.

<!--
This is where most fisheries biologists will probably stop reading.
Modern stock assessment models frequently have over 100 parameters.
However, the goal of exercise understand how a good forecast can lead to a bad decision even when we use these iterative forecasting methods correctly, not predict the safe harvest levels of Pacific Anchovy in 2022.
If any method fails to select the better model in the simplest cases, it does not often do better in more complex ones, even though the problems are harder to detect.
-->

Each possible parameterization of the model will generate a different forecast, and thus require a it's own SDP solution to determine the corresponding policy.
We will consider a larger suite of these logistic models in a moment, but first let us draw out only two such parameterizations from this larger suite.
This will let us focus our attention on two of the most interesting regions of that are already included within that larger parameter space of all possible values for $r$, $K$, and $\sigma$.
Thus, let us take "Model 1" as being given by $r_1 = 2$, $K_1 = 16$, $\sigma_1 = 0.05$, "Model 2" by $r_2 = 0.5$, $K_2 = 10$, $\sigma_2 = 0.075$. 
We can imagine our comparison of these two models as a microcosm of the larger comparison between all possible paremeterizations.

Ecologists will rightly scoff at the simplicity of these models -- the real world is much more complicated.
So it is important to bear in mind that these are not models that seek to approximate the stock dynamics of real world fisheries, only to approximate whatever "true model" we are using to drive the simulation.
In recognition of the fact that real world is always more complex than even our best ecological models, we will assume a "true model" for the simulations that is not in the Gordon-Schaefer class (i.e. our candidate models will never contain the true model), but is not so rich that a Gordon-Schaefer curve would seem a hopelessly poor approximation.

For illustrative purposes, we will thus assume the "true" process to be given by Model 3, which is unknown to the decision-maker, but similar enough to at least one of the candidate models might be considered a reasonable approximation:

\begin{equation}
f_3(Y) = Y + r_3 Y^4 \left( 1 - \frac{Y}{K_3} \right)
\end{equation}

with $r_3 = 0.002$, $K_3 = 10$ and $\sigma_3 = 0.05$.

Certainly, the challenge of choosing which model to base a decision policy on in the real world is much harder than this binary choice between two models, and yet it is sufficient to illustrate the trap. 
We will see later why making the models much more complex does not guarantee that the task becomes easier or that the trap may be ruled out.

# Methods for Managing Under Model Uncertainty

I will use this example to illustrate two alternative approaches for iterative learning over model uncertainty: iterative forecasting and adaptive management.
The central difference in the approaches is that iterative forecasting is premised on the ability to score the predictions of alternative models.
Iterative forecasting is silent on the issue of what to do with those scores, this is left up to the decision-maker.
Adaptive management approaches, by contrast, explicitly seek to integrate probabilities over all candidate models to reach a decision.
I consider each in turn.

## Statistical approaches: Forecasting under "Proper" Scoring Rules

Like many decision problems, the task of setting a sustainable harvest quota appears to hinge on having an accurate forecast: if we can predict to what size the fish stock will increase next year, $X_t+1$, and we know the current stock, $X_t$, then we can sustainably harvest $X_{t+1} - X_t$ without decreasing the biomass over the long term.
Selecting a model based on forecast skill is also justifiable on theoretical grounds, since it reduces the risk of overfitting by comparing model predictions to later observations that were not used to estimate the model [@Gneiting2014].

I illustrate the process of model selection by strictly proper scoring rules using two scenarios.
In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under model 1 and model 2 respectively [Fig 1].
The mean, $\mu_t$ and variance, $\sigma_t$ of the forecast are compared against the true observation $x_t$ using a proper scoring rule given by @Gneiting2007,

\begin{equation}
S(x_t|\mu_t,\sigma_t) = -(\mu_t - x_t )^2 / \sigma_t^2 - \log(\sigma_t) \label{proper}
\end{equation}

for each prediction over 100 replicate simulations of 100 time steps each [Fig 1].

```{r figure1, fig.cap = "Forecast performance of each model. Panels A, B: Step ahead predictions of stock size under unfished (A) and fished (B) scenarios. Error bars indicating the 95% confidence intervals around each prediction, while stars denote the observed value in that year. Because the models make different decisions each year in the fished scenario, the observed stock size in year 2, 3, etc under the management of model 1 (blue stars) is different from that under model 2 (red stars). Panels C, D: corresponding distribution of proper scores across all predictions (100 replicates of 100 timesteps). Higher scores are better, confirming that model 2 makes the better forecasts."}
predictions <- read_csv("../data/predictions.csv") %>% mutate(model = as.character(model))

fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C.",
    B_fished = "D."))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

fig1ab <- predictions %>% 
  filter(reps == "2", time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, 
             alpha = 0.6) + 
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "A. Without fishing",
    B_fished = "B. Managed harvest"))) +
  ylab("stock size")

fig1ab + fig1cd + 
  plot_layout(widths = c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

In Scenario B (actively harvest), I have first solved for the optimal management strategy for both model 1 and model 2 [Fig 1b] using SDP, [see @Marescot2013 or code in the Appendix].
Under this scenario, replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either model 1 and model 2.
The resulting stock sizes in the time-step following this harvest are once again compared to the probabilities predicted by each model using Eq \eqref{proper}.
Model 2 unequivocally outperforms model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of model 2 in both scenarios, the outcomes from management under model 2 are substantially worse.
We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass).
In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.
This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests.
While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig 2].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under model 1 are also substantially higher (panel B)"}

sims <- read_csv("../data/sims.csv")
npv_df <- read_csv("../data/npv_df.csv")

plot_ecology <- 
  sims %>%
  mutate(model = as.character(model), resp = as.character(reps)) %>%
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
  geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  labs(subtitle = "A. Ecological outcomes")

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

plot_econ <- 
npv_df %>%
  mutate(model = as.character(model), resp = as.character(reps)) %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time") + 
  labs(subtitle = "B. Economic outcomes")

plot_ecology + plot_econ
```

In both scenarios, the careful comparison through proper scoring rules has led us to select the worse-performing model.
Crucially, a manager operating under this selection would have little indication that their model was flawed: both future stock sizes and expected harvest yields consistently match model predictions.
Had we been able to include Model 3 in our forecast comparisons, it would equal or outperform the forecasting skill of both model 1 and model 2 (as guaranteed by the theorem of @Gneiting2007), while also matching or out-performing their economic utility (as guaranteed by the theorem of @Reed1979).
In practice, we never have access to the generating model, so it is reasonable to expect model selection to determine the best approximation.
As we see here, the best approximation for forecasting future states does not in fact lead to better outcomes.

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the simulation.
It would be possible to generate forecasts for the next $n$ steps, but these forecasts would also be conditional on the management action (i.e. harvest quota) selected at each step.
To evaluate two-step-ahead predictions we must consider each possible action under each possible state predicted by the step-ahead forecast, weighted by probability of the state given the model and the probability of the model.
This rapidly expanding set of possibilities is addressed by adaptive management for sequential decision problems [e.g. @Smith1981], which I employ in the next section.

## Decision-Theoretic Approaches

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982].
Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.
I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.
Passive adaptive management for a simple sequential decision problem is straightforward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix].
To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.
After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct [Fig 3A].
I compare these results to 
As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating, which achieves a net present value that only 31% that expected under management using model 1 alone [Fig 2].

```{r figure3, fig.cap="Adaptive management under model uncertainty. Solid lines trace the trajectories of the state (fish stock, circles) and action (harvest quota, triangles), under adaptive management (learning). Dotted lines trace the corresponding trajectories if iterative learning is omitted, leaving the prior belief fixed throughout the simulation (planning). Color indicates the belief that model 1 is correct (blue), with an initial prior belief of 99%. Panel A: Management over the two candidate models, Model 1 and Model 2. Within a single iteration of adaptive management, the belief over models switches from a prior belief that heavily favored model 1 to a posterior that favors model 2 with near certainty. Future iterations reinforce the belief in model 2, resulting in both depressed harvests and low stock sizes (solid lines). If no iterative learning updates are performed, stock sizes and realized harvests (and thus economic profit) are both higher. Panel B: given 42 candidate models over a broad range of parameter values, adaptive management quickly reduces the probability of model 1, and substantially underperforms management without learning (dotted lines). While outcomes improve marginally relative to the two-model case (figure A) they remain significantly worse than had no iterative learning been included."}
am <- read_csv("../data/am.csv")
fig3a <- am %>% filter(time <=20) %>% 
  pivot_longer(c(stock, quota), values_to="stock", names_to="key") %>%
  mutate(key = factor(key, levels = c("stock", "quota"))) %>%
  ggplot(aes(time, stock, shape = key, group = interaction(method,key))) + 
  geom_line(aes(lty=method), col = "gray40", show.legend = FALSE) + 
  geom_point(aes(col = belief), size=2, show.legend = FALSE) +
  scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
  ylab("fish stock") + labs(subtitle="A.")

am_multi <- read_csv("../data/am_multi.csv")
fig3b <- am_multi %>% 
  pivot_longer(c(stock, quota), values_to="stock", names_to="key") %>%
  mutate(key = factor(key, levels = c("stock", "quota"))) %>%
  ggplot(aes(time, stock, shape = key, group = interaction(method,key))) + 
    geom_line(aes(lty=method), col = "gray40") + 
    geom_point(aes(col = belief), size=2) +
    scale_colour_gradient(limits = c(0,1), low = pal[4], high = pal[1]) +
    ylab("fish stock") + labs(subtitle="B.")

fig3a + fig3b
```

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does little to resolve this problem [Fig 3B]. 
Iterative updates again quickly dismiss the parameter values assumed by model 1, though with more options to choose from, this probability is spread over a range of seemingly plausible candidate models instead of a single alternative model (see Appendix).
While the adaptive management of additional actions and observations slowly narrow this subset of plausible models, decisions based on this uncertainty prevent fish stocks from recovering fully, and realize lower harvests as a result.
Note that learning under either adaptive management approach (using two models or 42), the decision-maker becomes ever-increasingly convinced that they are using the right model or models.
Future stock sizes fall consistently in the range predicted by the model(s), and consistently outside the range predicted by model 1.
Consequently, each iteration the managers are only more firmly convinced that they are maintaining the fish stock near the biomass that supports maximum sustainable yield, when in fact they are sustaining a harvest regime that is preventing recovery of the stock to the much higher productivity regime which would have been achieved under model 1.

# Discussion

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other.
Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.
In both approaches, the manager quickly concludes that model 1 is entirely implausible, while finding that model 2 is remarkably accurate at predicting future values -- the manager is unknowing trapped into bad outcomes.
This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models spanning a wide range of possible $K$ and $r$ values): which improves only marginally over the outcome of considering only two models [Fig 3B].
Without access to the true model, none of our methods for model choice or reducing model uncertainty can break this trap.
This should not be mistaken as a critique of adaptive management or iterative forecasting specifically -- other model choice approaches such as goodness-of-fit, information criteria or cross-validation would all prefer model 2 as well.

<!-- Intuition, Figure 4 -->
The reason for model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, model 3.
Plotting the growth rate functions of each model, [Fig 4A], it is hardly surprising that no method exists which would not prefer the closely overlapping model 2 to the no-where-close model 1 as the better approximation of model 3.
Nevertheless, decisions based on model 1 are nearly indistinguishable from those based on the true model [Fig 4B], while model 2 leads to over-harvesting.
The explanation comes from noticing that the stock size corresponding to the maximum growth rate under model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for model 3.
Meanwhile, the peak of model 2 occurs at a substantially lower stock size.
While the optimal control solution appears to depend only on a step-ahead forecast accuracy (recall that the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.[^1]
Consequently, many models can share the same optimal solution, regardless of their forecast performance.

This result may appear surprising, and begs the question as to whether the non-uniqueness of the optimal solution is not merely an artifact of the simple models considered here. 
Surely more any realistically complex model would require a comparably more complex management policy such that the likelihood that any of our candidate models shares the same optimal policy would be vanishingly small?
The intuition sounds reasonable

[^1]: This is only approximately accurate. @Reed1979 demonstrates

This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

<!-- Intuition for it's not just this example -->
Situations similar to this example, in which a model can lead to the optimal decision while producing a poor forecast, are likely to be widespread.

<!--
For example, marine fisheries typically restrict a Management Strategy Evaluation [MSE; @Punt2016] to constant mortality, (defined as harvest per unit biomass, $F = H/B$).
Constant mortality is technically only optimal in a deterministic model [@Clark1973]; but is easier to solve for than the SDP methods for constant escapement shown here,
which in practice is used mostly in salmon fisheries and the resource economics literature [see @pomdp-intro, @Costello2016].
While Model 1 used here will not lead to a good constant-mortality policy, it is just as easy to construct a model which gives optimal decisions and bad forecasts.
In the example shown above, Model 1 achieves optimal results by choosing $K$ such that $K/2$ matches the peak growth rate under the true model [@Reed1979.
Under constant mortality, we instead choose a logistic model with $r$ such that $r/4$ matches the optimal constant mortality under the true model [@Schaefer1954].
Everything else about our choice of Model 1 can be far from the true model, and yet it will give optimal results, just as above.
-->

```{r figure4, fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth. Panel B: The optimal control policy under Model 1 is nearly identical to that under the true Model 3, while the optimal policy under Model 2 suppresses stock to a much lower escapement level."}
# Fig 4
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

plot_models <- 
  model_curves %>% 
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")

plot_policies <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")

plot_models + labs(subtitle="A") + plot_policies + labs(subtitle="B")
```

This simple example can be resolved by including our true model in the candidate set, or any class of models flexible enough to 'span' the true model, such as a Gaussian Process [e.g. see @Boettiger2015].  
Given sufficient data, either approach shown here will prefer the true model over any other (as guaranteed by @Gneiting2007's theorem), and thus discover the optimal policy.
While it is not difficult to image a candidate set of models flexible enough to include the very simple "true model" from Eq (4), it is much more unlikely that our model candidates truly span the generative processes of the real world.

This example demonstrates that it is possible for a model to give quite good forecasts while leading to marginal outcomes, and conversely, that is possible for a model to give optimal outcomes while making terrible forecasts.
This result may be particularly surprising given that the policy is derived directly from the forecast.

In most realistic management scenarios, we do not have the benefit of a theorem similar to Reed's to identify what features are most critical to approximate well and what are not.

For instance, the theorem no longer holds if stock estimates are uncertain [@pomdp-intro].
Absent the guidance of such results, iterative learning over even relatively rich possible candidate models can move us towards worse outcomes than not learning at all.

That potential for long-term deterioration in management outcomes due to adaptive learning has not been widely acknowledged before.
Certainly the limitations of any modeling approach to improve management outcomes due to the human context of competing interests, imperfect enforcement, etc., is readily appreciated [@Ludwig1993].
The problem here goes deeper, to the limits of models and learning from data even in a toy example free from other real world challenges.
Previous work has acknowledged that iterative learning may not always be feasible and that it may not always be beneficial [@Polasky2011], but has failed to recognize the possibility that it can potentially be detrimental.
For example, active adaptive management is premised on the observation that future observations may be too uninformative to distinguish between alternatives [@Walters1978].
Others have also noted that even big improvements in predictive accuracy could have negligible improvement on the decision outcomes, which motivates quantifying the Value of Information (VOI) [@Katz1987].
But in both cases, the worst outcome is wasted effort.
In the scenario considered here, the VOI is negative: any learning over the proposed model uncertainty leads to lower expected net utility than not learning.

Worse, absent the counter-factual of non-adaptive management to compare to, this deterioration in outcomes is invisible. 
All the information available to our manager reinforces the conviction that this deteriorated ecological state is in fact the natural, sustainably managed equilibrium of our fish population.
By responding in any way to future observations to reduce model uncertainty, our manager is caught in a trap of self-fulfilling prophecy:
believing ever more firmly in models that make more accurate predictions while driving decisions that prevent the fish stock from recovering.
The only way to avoid this trap is to _forego learning_ with each new observation, to avoid iterative updates.
It may seem that the adaptive management approach fails because posterior probabilities are updated according to Bayes rule.
Like iterative forecasting, this favors models which better predict the data, despite the fact that under the adaptive management approach, the overall optimization is conditioned on actual utility and not model fit.
However, this issue is not easily avoided.
For example, so-called greedy optimization techniques that favor actions with higher immediate reward do even worse in this context, since such an algorithm would obviously maximize the immediate harvest and therefore collapse the stock.
Only by comparing the net utility derived from management under model 1 for many iterations to the net utility derived under model 2 after many iterations can we determine that model 1 leads to better outcomes.
This raises several questions.
Do such methods of learning and reducing model uncertainty ever lead us towards worse long-term outcomes in the real world? 
If so, How would we know, and what should we do about it?

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?
It is tempting to argue that model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.
Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [e.g. @Tetlock2015].
If there is no shortage of bad forecasts, then could any of them be useful? 
The premise that a model need not perfectly capture reality to be useful is at the very heart of modeling. 
It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].
This makes it more likely that such models could persist in practice until now. 
While iterations that revise these models will no doubt improve decision outcomes in many cases, it is worth bearing in mind from the example here that such improvement is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them?
I believe this is an open question, but that our first step must be to recognize it as such.
We have seen that the problem cannot be resolved by more data alone, and is not the result of over-fitting.
Introducing more models into our candidate set is only helpful to the extent that such models can capture features important to the decision, which as we have seen can be quite different from the feature which matter most to prediction.
We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations.
Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, and essential strategies to avoid the very real problem of model over-fitting.
But their very success in selecting models which give the most accurate predictions can also be trap.
How that is best done in general is an open question.

## Acknowledgements

The author acknowledges support from NSF CAREER Award #1942280 and helpful discussions with Melissa Chapman, Jeremy Fox, and anonymous reviewers.

\pagebreak

# References