### 2022-02-04 notebook/alternate-models.Rmd
WIP :construction: alternate examples

---
output: github_document
---

- Define a model
- Define action space. Under the assumption of fixed action, problem is not MDP and trivial to solve by sweep through possible fixed actions.

```{r}
library(tidyverse)
```

## Five species

```{r}

model_5sp <- function(u_prey = 0.5,
                      u_pred = 0.5,
                      sigma = 0.05,
                      x1o = 0.44, # unfushed equib
                      x2o = 0.17,
                      x3o = 0.186,
                      x4o = 0.600,
                      x5o = 0.280,
                      Tmax = 50) {
    x1 <- x2 <- x3 <- x4 <- x5 <- numeric(Tmax)
    s1 <- s2 <- s3 <- s4 <- numeric(Tmax)
    xi1 <- rnorm(Tmax, 0, sigma)
    xi2 <- rnorm(Tmax, 0, sigma)
    xi3 <- rnorm(Tmax, 0, sigma)
    xi4 <- rnorm(Tmax, 0, sigma)
    xi5 <- rnorm(Tmax, 0, sigma)

    x1[1] <- x1o
    x2[1] <- x2o
    x3[1] <- x3o
    x4[1] <- x4o
    x5[1] <- x5o

    for (t in 1:(Tmax-1)) {
        s1[t] <- x1[t] * (1 - u_prey)
        s2[t] <- x2[t] * (1 - u_prey)
        s3[t] <- x3[t] * (1 - u_prey)
        s4[t] <- x4[t] * (1 - u_pred)

        x1[t+1] = s1[t]*exp(1.0213 - s1[t] - 0.0861*s2[t] - 0.3141*s3[t] - 0.7252*s4[t] - 0.2445*x5[t] + xi1[t])
        x2[t+1] = s2[t]*exp(1.0289 - 0.4765*s1[t] - s2[t] - 0.1370*s3[t] - 0.9811*s4[t] - 0.0915*x5[t] + xi2[t])
        x3[t+1] = s3[t]*exp(1.0207 - 0.3193*s1[t] - 0.3461*s2[t] - s3[t] - 0.6367*s4[t] - 0.8716*x5[t] + xi3[t])
        x4[t+1] = s4[t]*exp(0.7252*s1[t] + 0.9811*s2[t] + 0.6367*s3[t] - s4[t] + xi4[t])
        x5[t+1] = x5[t]*exp(0.2445*s1[t] + 0.0915*s2[t] + 0.8716*s3[t] - x5[t] +  xi5[t])

    }
    df <- tibble(time = 1:Tmax, x1, x2, x3, x4, x5)
    df
}

```

```{r}
clip <- function(x, lower=0, upper=1) {
    x[x <= lower] <- lower
    x[x >= upper] <- upper
    x
}
```

```{r}

reward <- function(df,
                   x,
                   w3 = 0.5,
                   w2 = 0.25,
                   w1 = 1 - (w2+w3),
                   delta = 0.999
                  ) {
    u_prey <- clip(x[1], 0, 1)
    u_pred <- clip(x[2], 0, 1)
    R <-
        w1 * u_prey * (df$x1 + df$x2 + df$x3) +
        w2 * u_pred * df$x4 +
        w3 * df$x5
    t <- seq_along(R)
    as.numeric(R %*% delta ^ t)
}

utility <- function(x, f = model_5sp) {

    u_prey <- clip(x[1], 0, 1)
    u_pred <- clip(x[2], 0, 1)
    ## ought to average over many replicates!
    df <- f(u_prey, u_pred)

    # negative since optimizer minimizes
    - reward(df, x)
}

```

```{r}
-utility(c(0.5,0.5))
-utility(c(0,1))
-utility(c(1,0))
-utility(c(0,0))

o <- optim(c(0.,0.), utility, method="L-BFGS-B", lower = c(0,0), upper = c(1,1))
o$par
o$value

o <- optim(c(0.5,0.1), utility)
clip(o$par, 0, 1)
o$value

#o <- optim(c(0.5,0.5), utility, method="SANN")
#clip(o$par, 0, 1)
#o$value

```

For the sake of argument, let us allow the model of Brias & Munch to represent the "true" model, which is unknown.
As researchers, we only have access to data derived from observations from the model. 
Let us consider a sample simulation of such data:

```{r}
df <- model_5sp(0, .6, sigma = 0.01)

df %>%
    pivot_longer(tidyselect::starts_with("x"),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(time, biomass, col= species)) + geom_path()
```

There are many other models, including process-based models, statistical models or even black-box machine learning models
which 5 dimensional state-space and two-dimensional action space.

```{r}
## Three-species 'view' of the underlying dynamics, unfished

df <- model_5sp(0, 0., sigma = 0.01, x1o =.1, x2o = .1, x3o = .1, x4o = .1, x5o = .1)
df %>%
    transmute(time = time, cormorant = x5, bass = x4, herring = x1+x2+x3) %>%
    pivot_longer(any_of(c("cormorant", "bass", "herring")),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(time, biomass, col= species)) + geom_path()
```

Now consider an alternative model which ignores some of this biological complexity:
Our alternate model has several oversimplifications:
(1) it treat all three herring species as a single species,
(2) it fails to capture the coupled predator-prey dynamics between bass and herring, 
(3) it oversimplifies the cormorant dynamics, assuming the cormorant population is determined to be a fixed fraction of the herring,
and ignoring the impact of cormorant's predation on the herring itself.
(4) Lastly, our model will overestimate the mortality introduced on herring by a given fishing effort.
These elements are all obviously wrong, but not so arbitrary as to be inconceivable as a candidate model.
Researchers frequently consider models which make oversimplifications all the time, and rely on model choice processes to weed them out.

The parameterization chosen for the oversimplified 3 species model can reasonably reproduce the unfished dynamics.
Under any harvesting regime or other influence that perturbs the system significantly far from the unfished equilibrium co-existence state of the model quickly reveals the poor forecasting ability of this model.
Nevertheless, solving for the optimal policy (under the same constraints of constant harvest fractions as we consider for the true model) results in a harvest policy which provides nearly optimal performance.

```{r}
model_3sp <- function(u_prey = 0,
                      u_predator = 0,
                      A1 = 2,
                      R1 = 1.6,
                      A2 = 2.75,
                      R2 = 1.6,
                      Tmax = 50,
                      x1o = .8, # Herring
                      x2o = .6, # Bass
                      x3o = 0.15, # Cormorant
                      sigma = 0.01) {

    omega1 <- rnorm(Tmax, 0, sigma)
    omega2 <- rnorm(Tmax, 0, sigma)
    omega3 <- rnorm(Tmax, 0, sigma)

    s1 <- x1 <- numeric(Tmax)
    s2 <- x2 <- numeric(Tmax)
    x3 <- numeric(Tmax)
    x1[1] <- x1o
    x2[1] <- x2o
    x3[1] <- x3o

    for (t in 1:(Tmax-1)) {
        s1[t] <- x1[t] * (1 - 10*u_prey)
        s2[t] <- x2[t] * (1 - u_predator)

        x1[t+1] <- s1[t] * exp(R1 - s1[t] * A1 + omega1[t])
        x2[t+1] <- s2[t] * exp(R2 - s2[t] * A2 + omega2[t])
        x3[t+1] <- .25 * x1[t]
    }

    df <- tibble(time = 1:Tmax, x1, x4 = x2, x5 = x3, x2 = 0, x3 = 0)
}
```

```{r}
# unfished equilibrium
df <- model_3sp(0, 0)
df %>%
    transmute(time = time, cormorant = x5, bass = x4, herring = x1+x2+x3) %>%
    pivot_longer(any_of(c("cormorant", "bass", "herring")),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(time, biomass, col= species)) + geom_path()
```

```{r}
# Optimal policy under identical utility
u_3sp <- function(x) utility(x, f = model_3sp)

m3sp <- optim(c(0.5,0.1), u_3sp)
m3sp$value
clip(m3sp$par, 0, 1)

```

Alternate model, better forecast, worse decision:

```{r}
model_B <- function(u_prey = 0.5,
                      u_pred = 0.5,
                      sigma = 0.05,
                      x1o = 0.796, # unfushed equib
                      x4o = 0.600,
                      x5o = 0.280,
                      Tmax = 50,
                      R1 = 1.0213,
                      A11 = 1,
                      A14 = 0.7252,
                      A15 = 0.2445,
                      A41 = 0.7252,
                      A51 = 0.2445) {
    x1 <- x4 <- x5 <- numeric(Tmax)
    s1 <-  s4 <- numeric(Tmax)
    xi1 <- rnorm(Tmax, 0, sigma)
    xi4 <- rnorm(Tmax, 0, sigma)
    xi5 <- rnorm(Tmax, 0, sigma)

    x1[1] <- x1o
    x4[1] <- x4o
    x5[1] <- x5o

    for (t in 1:(Tmax-1)) {
        s1[t] <- x1[t] * (1 - u_prey)
        s4[t] <- x4[t] * (1 - u_pred)
        x1[t+1] = s1[t]*exp(R1 - A11*s1[t] - A14*s4[t] - A15 * x5[t] + xi1[t])
        x4[t+1] = s4[t]*exp(A41*s1[t] - s4[t] + xi4[t])
        x5[t+1] = x5[t]*exp(A51*s1[t] - x5[t] +  xi5[t])

    }
    df <- tibble(time = 1:Tmax, x1, x4, x5, x2 =0, x3 = 0)
    df
}
```

```{r}
# Optimal policy under identical utility
u_B <- function(x) utility(x, f = model_B)
u_B(c(0, .63))
#optim(c(0.5,0.1), u_B, method = "L-BFGS-B", lower=0, upper=1)

B <- optim(c(0.5,0.1), u_B)
B$value
clip(B$par, 0, 1)

```

```{r}

utility( c(clip(o$par[[1]]), clip(o$par[[2]])) )

utility( c(clip(m3sp$par[[1]]), clip(m3sp$par[[2]])) )

utility( c(clip(B$par[[1]]), clip(B$par[[2]])) )

```

Run the simulation using management policy of each

```{r}

## Generate replicate predictions from each model at each step.
## compute CRPS of an observation from the "true" model vs those distributions

function(est, f) {
    u_prey <- clip(est$par[[1]])
    u_pred <- clip(est$par[[2]])
    map_dfr(1:100, function(i) {
    df <- f(u_prey, u_pred, x1o = .1, x2o = .1, x3o = .1, x4o = .1, x5o = .1)
    df %>%
        transmute(time = time, cormorant = x5, bass = x4, herring = x1+x2+x3) %>%
        pivot_longer(any_of(c("cormorant", "bass", "herring")),
                        names_to = "species",
                        values_to = "biomass")
    }, .id = "rep")

}

```

```{r}
# management of true model under optimal pars for model_3sp
df <- model_5sp(clip(m3sp$par[[1]],0,1),
                clip(m3sp$par[[2]],0,1),
                 x1o = .1, x4o = .1, x5o = .1)
df %>%
    transmute(time = time, cormorant = x5, bass = x4, herring = x1+x2+x3) %>%
    pivot_longer(any_of(c("cormorant", "bass", "herring")),
                    names_to = "species",
                    values_to = "biomass") %>%
    ggplot(aes(time, biomass, col= species)) + geom_path()
```